---
title: "MSA 6701 Project"
author: "Nick Wawee"
date: "11/12/2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
rm(list=ls())
library(MASS)
library(StepReg)
library(car)
library(ggplot2)
library(ggpubr)
library(reshape2)
library(limma)
library(ggrepel)
knitr::opts_chunk$set(echo = TRUE)
plot_opts =   theme_bw()+
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black") , axis.title.x = element_text(face='bold', size = 14), axis.title.y = element_text(face='bold', size = 14), plot.title = element_text(face='bold', size = 18, hjust=0.5))
```

```{r functions, include = F}
plotdists = function(df, path, brtype = 'Scott'){
  plotlist = list()
  for (col in colnames(df)){
    x = df[,col]
    if (class(x)=='factor'){
      dfnew = data.frame(col = factor(), count = integer())
      for (level in levels(x)){
        count = length(which(x==level))
        dfnew = rbind(data.frame(col= level, count = count), dfnew)
      }
      dfnew$col <- factor(dfnew$col, levels = dfnew$col[order(dfnew$count)])
      p = ggplot(dfnew, aes(x=col, y=count))+
        geom_bar(stat= 'identity')+
        plot_opts+
        labs(x = col, y = 'Count')+
        geom_text(aes(label = count), vjust = -0.3) + theme(axis.title = element_text(size = 24))
      plotlist[[col]] = p
    }else if (class(x) == 'numeric' | class(x) == 'integer'){
      dfnew =data.frame(col = class(x))
      histinfo = hist(x = x , breaks=brtype, plot = F)
      p = ggplot(as.data.frame(x), aes(x=x))+
        geom_histogram(bins = length(histinfo$breaks))+plot_opts+
        geom_vline(aes(xintercept = median(x)), 
             linetype = "dashed", size = 1, col = 'red')+
        labs(x = col, y = 'Count')+ theme(axis.title = element_text(size = 24))
        
      plotlist[[col]] = p 
    }
  }
  pfinal =ggarrange(plotlist = plotlist)
  ggsave(paste(path,'dists.pdf',sep=""), pfinal, device = 'pdf',height=1.3*length(plotlist) , width=1.3*length(plotlist)*1.5, units="in", limitsize = FALSE, dpi=900)
  return(pfinal)
}

Assumption_Check = function(l_m, outp){
  df = l_m[['model']]
  df$residuals = l_m[["residuals"]]
  df = df[,-which(colnames(df)==l_m[["call"]][["formula"]][[2]])]
  #Assumption 1 - linearity check
  a1pls = list()#assumption 1 plotlist
  for (col in colnames(df)[-which(colnames(df)=='residuals')]){
    dfnew = data.frame(x = df[,col], y = df$residuals)
    a1pls[[col]] = ggplot(data = dfnew, aes(x =x, y=y))+
        geom_jitter(size = 0.3)+plot_opts+labs(x = col, y = 'Residuals')+
      theme(axis.text.x = element_text(angle = 45, hjust = 1))+
      geom_hline(yintercept=0,linetype="dashed",color = "red", size=1)+theme(axis.title.x = element_text(size = 10))
  }
  p1 = ggarrange(plotlist= a1pls)
  p1 = annotate_figure(p1,text_grob("Linearity Checks", color = "red", face = "bold", size = 20))
  ggsave(filename = paste(outp,'linearity.png',sep=""), plot = p1, dpi = 600, height = length(a1pls), width = length(a1pls)*1.5, units = 'in')
  
  #Assumption 2 - Bias and Scedasticity 
  df2 = data.frame(Fit = l_m$fitted.values, Residuals = l_m$residuals)
  p2 = ggplot(data = df2, aes(x = Fit, y = Residuals))+
    geom_point()+plot_opts+geom_hline(yintercept=0,linetype="dashed",color = "red", size=1)+labs(title = 'Bias and Scedasticity\nCheck')+theme(plot.title = element_text(color = 'red'))
  ggsave(filename = paste(outp,'bias_sced.png',sep=""), height = 4, width = 4, units = 'in', plot = p2, dpi = 600)
  
  #Assumption 3 - Correlation in Errors
  a3pls = list()
  for (n in colnames(df[,-which(colnames(df)=='residuals')])){
    dfnew = data.frame(y = df$residuals[order(df[,n])], x = 1:nrow(df))
    a3pls[[n]] = ggplot(data = dfnew, aes(x = x, y = y))+geom_jitter(size = 0.3)+
      plot_opts+labs(title = paste('Sorted By:\n', n), x = 'Index', y='Residuals')+geom_hline(yintercept=0,linetype="dashed",color = "red", size=1)+theme(plot.title = element_text(size = 10))
  }
  p3 = ggarrange(plotlist = a3pls)
  p3 = annotate_figure(p3,text_grob("Error Independence Check", color = "red", face = "bold", size = 20))
  ggsave(filename = paste(outp,'inderror.png',sep=""), plot = p3, dpi = 600, height = length(a3pls), width = length(a3pls)*1.5, units = 'in')

  #Assumption 4 - Normality of Residuals
  shapres = shapiro.test(l_m$residuals)
  p4 = ggplot(df)+
    geom_qq(aes(sample = residuals))+geom_qq_line(aes(sample= residuals))+
    plot_opts+ labs(title = paste('Normality of Residuals\n', 'Shaprio Wilks Results:\n W = ', as.character(round(shapres$statistic,3)), ', p = ', as.character(round(shapres$p.value,5))), x = 'Theoretical Values', y = 'Sample Values')+ theme(plot.title = element_text(color = 'red'))
  ggsave(filename = paste(outp,'normres.png',sep=""), plot = p4, dpi = 600, height = 5, width = 5, units = 'in')
  
  pfinal = ggarrange(plotlist = list(p1, p2, p3, p4))
  pfinal= annotate_figure(pfinal,text_grob("Model Assumption Check", face = "bold", size = 26))
  ggsave(filename = paste(outp,'all_assum.png',sep=""), plot = pfinal, dpi = 600, width = 16, height = 16, units = 'in')
  return(pfinal)
}

infl_analysis = function(l_m){
  k = length(l_m$coefficients)-1
  n = length(l_m$fitted.values)
  row_num = 1:n
  #Leverage points
  hatdf = data.frame(Values = hatvalues(l_m), Row_Num = row_num, Type = rep('Hat Values', length(row_num)), Point_Type = rep('Leverage', length(row_num)), Bound1 = 2*(k+1)/n, Bound2 = 2*(k+1)/n)
  hatdf$Label = NA
  inds = which(hatvalues(l_m)>2*(k+1)/n)
  if(length(inds)!= 0){hatdf$Label[inds] = row_num[inds]}
  #Outliers
  instdf = data.frame(Values = rstandard(l_m), Row_Num = row_num, Type = rep('Internally Standardized Residuals', length(row_num)), Point_Type = rep('Outlier', length(row_num)), Bound1 = 3, Bound2 = -3)
  instdf$Label = NA
  inds = which(rstandard(l_m) > 3 | rstandard(l_m) < -3)
  if(length(inds)!=0){instdf$Label[inds] = row_num[inds]}
  
  extdf = data.frame(Values = rstudent(l_m), Row_Num = row_num, Type = rep('Externally Standardized Residuals', length(row_num)), Point_Type = rep('Outlier', length(row_num)), Bound1 = 3, Bound2 = -3)
  extdf$Label = NA
  inds = which(rstudent(l_m) > 3 | rstudent(l_m) < -3)
  if(length(inds)!=0){extdf$Label[inds] = row_num[inds]}
  
  #Influential
  dfitsdf = data.frame(Values = dffits(l_m), Row_Num = row_num, Type = rep('DEFFITS', length(row_num)),Point_Type = rep('Influential', length(row_num)), Bound1 = 2*sqrt((k+2)/(n-k-2)), Bound2 = -2*sqrt((k+2)/(n-k-2)))
  dfitsdf$Label = NA
  inds = which(dffits(l_m) > 2*sqrt((k+2)/(n-k-2)) | dffits(l_m) < -2*sqrt((k+2)/(n-k-2)))
  if(length(inds)!=0){dfitsdf$Label[inds] = row_num[inds]}
  
  cddf = data.frame(Values = cooks.distance(l_m), Row_Num = row_num, Type = rep("Cook's Distance", length(row_num)),Point_Type = rep('Influential', length(row_num)), Bound1 = 1, Bound2 = 1)
  cddf$Label = NA
  inds = cooks.distance(l_m) > 1
  if(length(inds)!=0){cddf$Label[inds] = row_num[inds]}
  
  cvdf = data.frame(Values = covratio(l_m), Row_Num = row_num, Type = rep("Covariance Ratio", length(row_num)),Point_Type = rep('Influential', length(row_num)), Bound1 = 1 + 3*(k+1)/n, Bound2 = 1 - 3*(k+1)/n)
  cvdf$Label = NA
  inds = covratio(l_m) > 1 + 3*(k+1)/n | covratio(l_m) < 1 - 3*(k+1)/n
  if(length(inds)!=0){cvdf$Label[inds] = row_num[inds]}
  
  ret_df = rbind(hatdf, instdf, extdf, dfitsdf, cddf, cvdf)
  return(ret_df)
}
```

## Loading and Inspecting

```{r loading}
data(Boston)
df = Boston
for (col in colnames(df)){
  cat("Number of Missing values in ", col,": ", as.character(length(which(is.na(df[,col])))), "\n")
}

str(df)
df = as.data.frame(df)
head(df)
```

## Model Fitting

```{r mlr1}
mlr = lm(crim~., data =df)
summary(mlr)
```

## Stepwise Regression

```{r stepwise}
sbi = stepwise(df, y = 'crim', selection = "backward", select = 'AIC')
sbi$variate
```

```{r mlr2}
mlr2 = lm(crim~ zn + nox + dis + rad + ptratio + black + lstat + medv, data =df )
summary(mlr2)
```

## Colinearity Check

```{r colinear}
vif(mlr2)
```

Correlation between regressors: 

```{r correaltions between regressors}
cor(df[,which(colnames(df)%in%sbi$variate)])
```

We see that the intercept is not statistically significant and it doesnt make any sense to have a negative crime rate, so we remove it.

## Power Transformation

```{r mlr4}
mlr4 = lm(crim~ 0 +  rad*dis + ptratio + black+medv+rad*nox , data = df)
summary(mlr4)
bc = boxcox(mlr4, data = df)
lambda = bc$x[which.max(bc$y)]
lambda
dfnew = df
dfnew$crim = df$crim^lambda

```

```{r boxcox}
mlr5 = lm(crim~   0 +rad*dis + ptratio + medv + nox, data = dfnew)
summary(mlr5)
```

## Distribution of Variables

```{r pltdists}
dfnew2 = dfnew[,c('crim','rad','dis', 'ptratio', 'medv', 'nox')]
colnames(dfnew2) = c('Crime Rate', 'Radial Highway Accessibility', 'Employment Distance', 'Pupil Teacher Ratio',"Median Value of Home (/$1000)", 'NOx Concentration (pptm)')

mlr6 = lm(`Crime Rate` ~ 0 +`Radial Highway Accessibility`+`Employment Distance` + `Pupil Teacher Ratio` + `Median Value of Home (/$1000)`  +`NOx Concentration (pptm)`, data = dfnew2)


outp = "../plots/"
distplot = plotdists(dfnew2,outp, brtype = 'FD')
distplot
```


## Scatter Plot of All Variables

```{r scatter}
df.m = melt(dfnew2,id.vars = 'Crime Rate')
colnames(df.m) = c('Crime Rate', 'Variable', 'Value')
#df.m$Variable = apply(strsplit2(df.m$Variable, split = "_")[,1:5], 1, paste, collapse = " ")
ggplot(data=df.m, aes(x=Value, y = `Crime Rate`))+
  geom_point(size = 0.25)+plot_opts+facet_wrap(~Variable, scales = 'free_x')+
  ylab(paste('[Crime Rate]^', as.character(round(lambda,4)), sep = ""))+theme(axis.text.x=element_text(angle =45, vjust = 0.7))

ggsave(filename = paste(outp,'scatter.png',sep=""), dpi = 600, width = 1.5*5, height = 5, units = 'in')
```

## Assumption Check

```{r assumption_check, fig.width = 15, fig.height = 15}
pf = Assumption_Check(mlr6, outp)
pf
```

## Influential Analysis

```{r infl_analysis, fig.height = 8, fig.width = 1.5*8}
ret_df = infl_analysis(mlr6)
ret_df = cbind(ret_df, dfnew2)
p = ggplot(data= ret_df, aes(x= Row_Num, y = Values))+
  geom_point(size = .8)+
  facet_wrap(~Type, scales = "free_y")+plot_opts+geom_line(aes(y=Bound1))+geom_line(aes(y=Bound2))+
  #geom_label_repel(aes(label=Label), size = 4)+
  labs(title = 'Influential Point Analysis', x = 'Observation Number')+ theme(strip.text = element_text(face = 'bold'))
p
ggsave(filename = '../plots/infl.png', plot = p, width = 1.5*10, height = 10, units = 'in', limitsize = F)
```

```{r infl_analysis_labeled, fig.height = 8, fig.width = 1.5*8}
p = ggplot(data= ret_df, aes(x= Row_Num, y = Values))+
  geom_point(size = .5)+
  facet_wrap(~Type, scales = "free_y")+plot_opts+geom_line(aes(y=Bound1))+geom_line(aes(y=Bound2))+
  #geom_label_repel(aes(label=Label), size = 4)+
  labs(title = 'Influential Point Analysis', x = 'Observation Number')+ theme(strip.text = element_text(face = 'bold'))
p
ggsave(filename = '../plots/infl.png', plot = p, width = 1.5*6, height = 6, units = 'in', limitsize = F)
```

## Hypothesis Testing

Test for significance of regression:
H0: All regression coefficients are equal to zero. H1: At least one regression coefficient is not equal to 0.

```{r sig_reg}
linearHypothesis(mlr5, c('rad = 0', 'dis = 0', 'ptratio = 0', 'medv = 0', 'nox = 0'))
```

## Confidence Intervals

```{r confin}
rmse = sqrt(mean(mlr6$residuals^2))
knitr::kable(confint(mlr6))
```

## Other Metrics

```{r other metrics}
rmse = sqrt(mean(mlr6$residuals^2))
rmse

AIC(mlr6)
BIC(mlr6)
```